Inverse energy transfer in three-dimensional quantum vortex flows

### 2025-06-26 main_REV.tex
GK done
Updated figures
Updated plots and andrews affiliation

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{blindtext}
\usepackage{float}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{xcolor}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\author{P. Z. Stasiak}
\author{A. Baggaley}
\author{C.F. Barenghi}
\affiliation{School of Mathematics, Statistics and Physics, Newcastle University, Newcastle upon Tyne, NE1 7RU, United Kingdom}

@@

\begin{figure*}
	\centering
    \begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snap-1.pdf}
	\end{subfigure}
	\begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snap-2.pdf}
	\end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snap-3.pdf}
	\end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snap-4.pdf}

	\end{subfigure} \hfill
    %
	\begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snaps-hel-1.pdf}
	\end{subfigure}
	\begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snaps-hel-2.pdf}
	\end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snaps-hel-3.pdf}
	\end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
		\centering
		\includegraphics*[width=\textwidth]{snaps-hel-4.pdf}
	\end{subfigure} \hfill
	\caption{
Three-dimensional rendering of the time evolution of
an initially orthogonal vortex configuration 
undergoing a vortex reconnection at $T = 1.9$K at dimensionless
times (from left to right)
$(t-t_R)/\tau_R=-7.47\times10^{-2}$, $-3.09\times10^{-3}$, $8.79\times10^{-4}$ and $5.39\times10^{-3}$, where $\tau_R = L_R/E^{1/2}_R$. The quantities $E_R$ and $L_R$ are the energy $E$ and integral length scales \red{$L$} of the normal fluid at reconnection time $t_R$. The green
tubes represent the superfluid vortex lines (the tubes’ radii have been 
greatly exaggerated for visual purpose). In the top sequence,
the blue volume rendering represents the scaled normal fluid enstrophy 
$\bom^2/\bom^2_{max}$. Note the Kelvin wave on the superfluid
vortex at $(t-t_R)/\tau_R=8.79\times10^{-4}$ . In the bottom sequence, the red/blue volume
rendering at the same times represent scaled positive/negative normal fluid
helicity.	
%Three-dimensional rendering of an orthogonal vortex configuration 
%undergoing a vortex reconnection at $T=1.9K$. The red tubes represent
%the superfluid vortex
%lines (the tubes' radii have been greatly exaggerated for visual purposes), 
%and the blue volume rendering represents the scaled normal fluid enstrophy 
%$\bom^2/\bom^2_{max}$. In the third panel, note the Kelvin waves on the superfluid vortices.
}
    \label{fig:visualisation}
\end{figure*}

@@

We now focus on the time evolution of the normal fluid energy spectrum $E(k)$, defined by 
\begin{equation}
    E = \frac{1}{(2\pi)^3}\int_{\mathcal{V}}\frac{1}{2}|\v_n|^2 dV = \int_0^{\infty}E(k)dk
\end{equation}
where $E$ is the total normal fluid energy and $k$ is the magnitude of the three-dimensional wavenumber. The energy at reconnection $E_R$ is given by $E(t_R)$, where $t_R$ is the time at reconnection. The energy spectrum $E(k)$ is displayed in Fig.~\ref{fig:kinetic-energy}.
\begin{figure}[h!]
    \centering
    \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics*[width=\textwidth]{energy-spec.pdf}
        \caption{}\label{fig:kinetic-energy}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics*[width=\textwidth]{flux-spec.pdf}
        \caption{}\label{fig:energy-flux}
    \end{subfigure}
    \caption{\emph{(a):} Normal fluid kinetic energy spectrum $E(k)$ before 
reconnection (dashed lines), at reconnection (solid lines) 
and after reconnection (dotted lines) for $T=1.9K$ (red)
and $T=2.1K$ (blue).\emph{(b):} Spectral normal fluid kinetic energy flux, $\Pi (k)$. It is normalised using the integral scale and the normal fluid energy at reconnection.
\emph{Inset:} Post reconnection evolution of the integral length scale, $L$.
Times and temperatures are labelled as in Fig.~\ref{fig:kinetic-energy}.}
\end{figure}
%

@@

% the dynamics of the normal fluid flow using the 
% spectral energy budget equation: 
% %
% \begin{equation}
%     \frac{\partial E(k)}{\partial t} = T(k) - D(k) + I(k)
% \end{equation}
% where $T(k)$ is the spectral kinetic energy transfer function,
% %(\textit{i.e.} $T(k)dk$ is the energy transfered per unit time by modes with $k < |\mathbf{k}| < k +dk$, determined by non-linear effects), 
% $D(k)=2\nu_n k^2 E_k$ is the dissipation spectrum and
% %=\mathrm{Re}(\hat{\mathbf{F}}_{ns}(\k)\cdot\hat{\v}^*(\k))$ 
% $I(k)$ is the injection spectrum arising from the mutual friction 
% force $\mathbf{F}_{ns}$.  
We observe that $\Pi(k) < 0$ for all $k$ during and after reconnection;
we also observe that, near the time of reconnection,
the peak value of $|\Pi(k)|$ is in the range $15 < k < 25$.
%
The negative sign of $\Pi(k)$ is evidence of a flux of kinetic  energy from small to large scales, exciting larger and larger scales. % In other words, at non-zero temperatures,
% vortex reconnections trigger an inverse transfer of energy, 
% which implies the creation of large scale structures, visible
% in Fig.~\ref{fig:visualisation}} \NOTE{GK: the figure shows small scale structures...}
This behaviour is quantified by the evolution of the integral length scale $L$, defined as
\begin{equation}
    L = \frac{\pi}{2 E}\int_0^{\infty}\frac{E(k)}{k}dk, %\Big/ \int_0^{\infty}\!\!E(k)dk.
\end{equation}
The inset of Fig.~\ref{fig:energy-flux} shows that
$L$ indeed  increases steadily in the post-reconnection regime. Note that times have been normalised by the largest eddy-turnover-time at the reconnection event, evidencing  it fast evolution.

To explain the inverse energy transfer shown in Fig.~\ref{fig:kinetic-energy},
we look whether the reconnection triggers a chirality imbalance. 
We decompose the incompressible Fourier modes of the normal fluid velocity 
into helical modes \cite{waleffe-1992}:
\begin{equation}
\hat{\mathbf{v}}_n (\k) = \hat{\mathbf{v}}_n^+(\k) +\hat{\mathbf{v}}_n^-(\k)=
 v_n^+(\mathbf{k}) \mathbf{h}^+(\mathbf{k})+v_n^-(\mathbf{k}) \mathbf{h}^-(\mathbf{k}),
\end{equation}
%
where $\mathbf{h}^\pm (\mathbf{k})$ are the two eigenvectors of the curl 
operator, \textit{i.e.} $i\k~\times~\h^{\pm}(\k)~=~\pm k \h^{\pm}(\k)$. 
Similarly, we decompose the transverse modes of the mutual friction force:
$\hat{\mathbf{F}}_{ns}^{\perp}(\k) = f^+(\k) \mathbf{h}^+ + f^-(\k) \mathbf{h}^-$
(the  Fourier modes of $\mathbf{F}_{ns}$ parallel to the wavemumber 
$\k$ do not play any role in the time evolution of $\mathbf{v}_n$ 
due to the incompressible constraint). Finally, the helical decomposition naturally allow us decompose the total helicity as $\mathcal{H}=\mathcal{H}^+-\mathcal{H}^-$ \cite{plunianInverseCascadeEnergy2020a}.
% %
% The spectral energy densities corresponding to the helical modes are
% $E^{\pm}(\k) = (1/2) |v_n^\pm(\k)|^2$, the total spectral density 
% being $E(\k) = E^{+}(\k) + E^{-}(\k)$. Similarly the spectral helicity 
% density is 
% \begin{equation}
% \begin{split}
% H(\k) & = (1/2) \hat{\mathbf{v}}_n (\k)\!\!\! ~\cdot~\!\!\! \hat{\bm{\omega}}_n^* (\k) = \\
% & = k E^+(\k) - k E^-(\k)  = H^+(\k) - H^-(\k),
% \end{split}
% \end{equation}
% where $\bm{\omega}_n$ is the normal fluid vorticity, the star indicates 
% complex conjugate and $H^\pm$ is the helicity contribution of each 
% separate helical mode. \NOTE{GK: the standard definition of the energy and helicity spectra contains the sum over the angles, so the previous definition are wrong (except if you redifined them for some reason).}
%The rate of change $I^\pm(\k)$ of the energies $E^{\pm}(\k)$ arising from the mutual friction is proportional to the corresponding helical coefficients $f^\pm$, \textit{i.e.} $I^\pm(\k) = \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$, and the related signed helicity injection is $I_H^\pm(\k) = k \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$.

@@

Our findings have profound implications for the nature of turbulence in 
finite temperature superfluids, \red{where vortex reconnections play a key role}. 
In circumstances where the vortex density 
$\mathcal{L}$ is large and where the isotropy of the vortex tangle 
is broken by external forcing (for instance thermal counterflows~\cite{vinen-1957a}), 
the chirality of the flow generated by the 
frequent \red{non-symmetrical} reconnections may be strong enough to induce an inverse energy 
cascade~\cite{notes_recon}.
%The asymmetry of the reconnecting vortex configurations is a key element; 
%when perfectly symmetrical vortex configurations (with corresponding 
%opposite vortex line orientations) reconnect, the injected helicity 
%globally cancels out, as verified in our study if we take also into 
%account the reconnection.
The microscopic mechanism that we have described \red{may be relevant in triggering}
the inverse energy cascade which is observed numerically in  
large-scale simulations of counterflow turbulence at large heat fluxes \cite{polanco2020}, which is
indeed inherently not isotropic.
%, hence triggering non-fully reconnections. 
Our work hence motivates further detailed studies of
the role played by helicity in superfluid dynamics 
\cite{dileoni2016,galantucci2021},
moving the emphasis from few vortex systems \cite{scheeler2014} 
to fully coupled superfluid turbulence.

>>>>

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{blindtext}
\usepackage{float}
\usepackage{caption,cancel}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{xcolor}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\author{P. Z. Stasiak}
\affiliation{School of Mathematics, Statistics and Physics, Newcastle University, Newcastle upon Tyne, NE1 7RU, United Kingdom}

\author{A. Baggaley}
\affiliation{School of Mathematics, Statistics and Physics, Newcastle University, Newcastle upon Tyne, NE1 7RU, United Kingdom}
\affiliation{Department of Mathematics and Statistics, Lancaster University, Lancaster, LA1 4YF, UK}

\author{C.F. Barenghi}
\affiliation{School of Mathematics, Statistics and Physics, Newcastle University, Newcastle upon Tyne, NE1 7RU, United Kingdom}

@@

\begin{figure*}
	\centering
  \includegraphics[width=\textwidth]{snapshots.pdf}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-4.pdf}

	% \end{subfigure} \hfill
  %   %
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-4.pdf}
	% \end{subfigure} \hfill
	\caption{
Three-dimensional rendering of the time evolution of
an initially orthogonal vortex configuration 
undergoing a vortex reconnection at $T = 1.9$K at dimensionless
times (from left to right)
$(t-t_R)/\tau_R=-7.47\times10^{-2}$, $-3.09\times10^{-3}$, $8.79\times10^{-4}$ and $5.39\times10^{-3}$, where $\tau_R = L_R/E^{1/2}_R$. The quantities $E_R$ and $L_R$ are the energy $E$ and integral length scales \red{$L$} of the normal fluid at reconnection time $t_R$. The green
tubes represent the superfluid vortex lines (the tubes’ radii have been 
greatly exaggerated for visual purpose). In the top sequence,
the blue volume rendering represents the scaled normal fluid enstrophy 
$\bom^2/\bom^2_{max}$. Note the Kelvin wave on the superfluid
vortex at $(t-t_R)/\tau_R=8.79\times10^{-4}$ . In the bottom sequence, the red/blue volume
rendering at the same times represent scaled positive/negative normal fluid
helicity.	
%Three-dimensional rendering of an orthogonal vortex configuration 
%undergoing a vortex reconnection at $T=1.9K$. The red tubes represent
%the superfluid vortex
%lines (the tubes' radii have been greatly exaggerated for visual purposes), 
%and the blue volume rendering represents the scaled normal fluid enstrophy 
%$\bom^2/\bom^2_{max}$. In the third panel, note the Kelvin waves on the superfluid vortices.
}
    \label{fig:visualisation}
\end{figure*}

@@

We now focus on the time evolution of the normal fluid energy spectrum $E(k)$, defined by 
\begin{equation}
    E = \frac{1}{(2\pi)^3}\int_{\mathcal{V}}\frac{1}{2}|\v_n|^2 dV = \int_0^{\infty}E(k)dk
\end{equation}
where $E$ is the total normal fluid energy and $k$ is the magnitude of the three-dimensional wavenumber. The energy at reconnection $E_R$ is given by $E(t_R)$, where $t_R$ is the time at reconnection. The energy spectrum $E(k)$ is displayed in Fig.~3a.
\begin{figure}
    \centering
    \includegraphics[width=0.49\textwidth]{energy-flux-spec.pdf}

    % \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{energy-spec.pdf}
    %     \caption{}\label{fig:kinetic-energy}
    % \end{subfigure}
    % \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{flux-spec.pdf}
    %     \caption{}\label{fig:energy-flux}
    % \end{subfigure}
% \begin{subcaptiongroup*}
% 	\phantomsubcaption\label{fig:minimum-distance}
% 	\phantomsubcaption\label{fig:prefactors}
% \end{subcaptiongroup*}

    \caption{\emph{(a):} Normal fluid kinetic energy spectrum $E(k)$ before 
reconnection (dashed lines), at reconnection (solid lines) 
and after reconnection (dotted lines) for $T=1.9K$ (red)
and $T=2.1K$ (blue).\emph{(b):} Spectral normal fluid kinetic energy flux, $\Pi (k)$. It is normalised using the integral scale and the normal fluid energy at reconnection.
\emph{Inset:} Post reconnection evolution of the integral length scale, $L$.
Times and temperatures are labelled as in Fig.~3a.}
\label{fig:spectra}
\end{figure}
%

@@

% the dynamics of the normal fluid flow using the 
% spectral energy budget equation: 
% %
% \begin{equation}
%     \frac{\partial E(k)}{\partial t} = T(k) - D(k) + I(k)
% \end{equation}
% where $T(k)$ is the spectral kinetic energy transfer function,
% %(\textit{i.e.} $T(k)dk$ is the energy transfered per unit time by modes with $k < |\mathbf{k}| < k +dk$, determined by non-linear effects), 
% $D(k)=2\nu_n k^2 E_k$ is the dissipation spectrum and
% %=\mathrm{Re}(\hat{\mathbf{F}}_{ns}(\k)\cdot\hat{\v}^*(\k))$ 
% $I(k)$ is the injection spectrum arising from the mutual friction 
% force $\mathbf{F}_{ns}$.  
We observe that $\Pi(k) < 0$ for all $k$ during and after reconnection;
we also observe that, near the time of reconnection,
the peak value of $|\Pi(k)|$ is in the range $15 < k < 25$.
%
The negative sign of $\Pi(k)$ is evidence of a flux of kinetic  energy from small to large scales, exciting larger and larger scales. % In other words, at non-zero temperatures,
% vortex reconnections trigger an inverse transfer of energy, 
% which implies the creation of large scale structures, visible
% in Fig.~\ref{fig:visualisation}} \NOTE{GK: the figure shows small scale structures...}
This behaviour is quantified by the evolution of the integral length scale $L$, defined as
\begin{equation}
    L = \frac{\pi}{2 E}\int_0^{\infty}\frac{E(k)}{k}dk, %\Big/ \int_0^{\infty}\!\!E(k)dk.
\end{equation}
The inset of Fig.~3b shows that
$L$ indeed  increases steadily in the post-reconnection regime. Note that times have been normalised by the largest eddy-turnover-time at the reconnection event, evidencing  it fast evolution.

To explain the inverse energy transfer shown in Fig.~3b,
we look whether the reconnection triggers a chirality imbalance. 
We decompose the incompressible Fourier modes of the normal fluid velocity 
into helical modes \cite{waleffe-1992}:
\begin{equation}
\hat{\mathbf{v}}_n (\k) = \hat{\mathbf{v}}_n^+(\k) +\hat{\mathbf{v}}_n^-(\k)=
 v_n^+(\mathbf{k}) \mathbf{h}^+(\mathbf{k})+v_n^-(\mathbf{k}) \mathbf{h}^-(\mathbf{k}),
\end{equation}
%
where $\mathbf{h}^\pm (\mathbf{k})$ are the two eigenvectors of the curl 
operator, \textit{i.e.} $i\k~\times~\h^{\pm}(\k)~=~\pm k \h^{\pm}(\k)$. 
Similarly, we decompose the transverse modes of the mutual friction force:
$\hat{\mathbf{F}}_{ns}^{\perp}(\k) = f^+(\k) \mathbf{h}^+ + f^-(\k) \mathbf{h}^-$
(the  Fourier modes of $\mathbf{F}_{ns}$ parallel to the wavemumber 
$\k$ do not play any role in the time evolution of $\mathbf{v}_n$ 
due to the incompressible constraint). Finally, the helical decomposition naturally allow us decompose the total helicity as $\mathcal{H}=\mathcal{H}^+-\mathcal{H}^-$ \cite{plunianInverseCascadeEnergy2020a}.
% %
% The spectral energy densities corresponding to the helical modes are
% $E^{\pm}(\k) = (1/2) |v_n^\pm(\k)|^2$, the total spectral density 
% being $E(\k) = E^{+}(\k) + E^{-}(\k)$. Similarly the spectral helicity 
% density is 
% \begin{equation}
% \begin{split}
% H(\k) & = (1/2) \hat{\mathbf{v}}_n (\k)\!\!\! ~\cdot~\!\!\! \hat{\bm{\omega}}_n^* (\k) = \\
% & = k E^+(\k) - k E^-(\k)  = H^+(\k) - H^-(\k),
% \end{split}
% \end{equation}
% where $\bm{\omega}_n$ is the normal fluid vorticity, the star indicates 
% complex conjugate and $H^\pm$ is the helicity contribution of each 
% separate helical mode. \NOTE{GK: the standard definition of the energy and helicity spectra contains the sum over the angles, so the previous definition are wrong (except if you redifined them for some reason).}
%The rate of change $I^\pm(\k)$ of the energies $E^{\pm}(\k)$ arising from the mutual friction is proportional to the corresponding helical coefficients $f^\pm$, \textit{i.e.} $I^\pm(\k) = \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$, and the related signed helicity injection is $I_H^\pm(\k) = k \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$.

@@

Our findings have profound implications for the nature of turbulence in 
finite temperature superfluids, \red{where vortex reconnections play a key role}. 
In circumstances where the vortex density 
$\mathcal{L}$ is large and where the isotropy of the vortex tangle 
is broken by external forcing (for instance thermal counterflows~\cite{vinen-1957a}), 
the chirality of the flow generated by the 
frequent \red{non-symmetrical} reconnections may be strong enough to induce an inverse energy cascade~\cite{notes_recon}.
%The asymmetry of the reconnecting vortex configurations is a key element; 
%when perfectly symmetrical vortex configurations (with corresponding 
%opposite vortex line orientations) reconnect, the injected helicity 
%globally cancels out, as verified in our study if we take also into 
%account the reconnection.
The microscopic mechanism that we have described \red{may be relevant in triggering}
the inverse energy cascade which is observed numerically in  
large-scale simulations of counterflow turbulence at large heat fluxes \cite{polanco2020}, which is indeed inherently not isotropic. \magenta{However, formalising the connection between high counterflow, unbalanced helicity injection and inverse energy cascade is  challenging, as the model used in \cite{polanco2020} and the one used in our work describe superfluid helium at very different scales.}
%, hence triggering non-fully reconnections. 
Our work hence motivates further detailed studies of
the role played by helicity in superfluid dynamics 
\cite{dileoni2016,galantucci2021},
moving the emphasis from few vortex systems \cite{scheeler2014} 
to fully coupled superfluid turbulence.
